---
title: "Demonstrating the Central Limit Theorem"
author: "Companion Package for Biostatistical Analysis of Proportions and Rates"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Central Limit Theorem Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

In this vignette, we demonstrate the Central Limit Theorem (CLT)
using the `cltdemo()` function. The CLT states that, regardless
of the original population distribution, the distribution of the
sample mean approaches a normal distribution as the sample size
increases.

Here, we use the Gamma distribution with varying skewness, which
is controlled by the `shape` parameter:

- **Shape = 0.5**: Highly skewed distribution
- **Shape = 1**: Exponential distribution (moderately skewed)
- **Shape = 2**: Less skewed distribution

We explore the behavior of the sample mean for different sample
sizes (`n = 5, 10, 20, 40`) to illustrate the convergence towards
the normal distribution.

# Load Required Packages

```{r setup, include = FALSE}
set.seed(123)
library("ibist")
```

# Gamma Distribution with Shape = 0.5

The Gamma distribution with `shape = 0.5` is highly skewed. We
expect to see the sample mean distribution becoming more normal
as the sample size increases.

```{r gamma-shape-0.5}
demo_clt(rgamma, n = c(5, 10, 20, 40), nrep = 10000,
         shape = 0.5, rate = 1, mean = 0.5, sd = sqrt(0.5))
```

# Gamma Distribution with Shape = 1

The Gamma distribution with `shape = 1` is equivalent to the
Exponential distribution. This example has moderate skewness,
and we observe the effect of increasing the sample size.

```{r gamma-shape-1}
demo_clt(rgamma, n = c(5, 10, 20, 40), nrep = 10000,
         shape = 1, rate = 1, mean = 1, sd = sqrt(1))
```

# Gamma Distribution with Shape = 2

The Gamma distribution with `shape = 2` has less skewness. Here,
the sample mean distribution converges more quickly to a normal
distribution even for smaller sample sizes.

```{r gamma-shape-2}
demo_clt(rgamma, n = c(5, 10, 20, 40), nrep = 10000,
         shape = 2, rate = 1, mean = 2, sd = sqrt(2))
```

# A General Distribution Constructed from a Mixture

When the population `mean` and `sd` are unspecified, the will be
approximated by a large sample (10,000) and then used in
standardization. For instance, consider sampling from the following
mixture distribution.

```{r mixture}
mymix_rng <- function(n, rate = 0.5) {
    ifelse(runif(n) < rate,
           rgamma(n, shape = 0.5), rgamma(n, shape = 4))
}

demo_clt(mymix_rng, n = c(5, 10, 20, 40))

```


# Conclusion

The plots above demonstrate the Central Limit Theorem in action.
As the sample size increases, the distribution of the sample mean
approaches a normal distribution, even for highly skewed
underlying distributions like the Gamma distribution with
`shape = 0.5`.

This vignette illustrates the robustness of the CLT and its
importance in statistical analysis, especially when dealing with
non-normal data in biostatistical contexts.
